Simulation Learning in Computer Hex

### 2011-11-04 thesis/thesis.tex
Added Fuego reference, and placeholder for content in Hivemind system detail section.
Working on the Hivemind description.

\begin{figure}
	\begin{HexBoard}[board size=13]
		\HGame{c4,f6,e11,i8,h8,i1,i6,d5,i3,
					 h9,h5,h4,i4,g9,f9,g8,f7,f8,e8}
	\end{HexBoard}
	\caption{13x13 Hex Game, \white{} to play.}
	\label{fig:13x13inprogress}
\end{figure}

@@

As stated, the current top computer Hex player, MoHex, uses the open-source Fuego implementation of Monte-Carlo Tree Search (MCTS). The basic Monte-Carlo technique for Computer Go was proposed by Bru\"{u}gmann\cite{brugmann1993monte}. His program \emph{Mango} effectively used a ``flat'' search tree (1-ply) and simply ran simulations from each candidate move to estimate its value. To contrast, MCTS uses Monte-Carlo simulations to grow a minimax search tree. The UCT algorithm was developed alongside MCTS to direct the growth of the search tree\cite{gelly2006exploration}. At any point, the search can be stopped and the current state of the tree used to select a move to play. Performance of the search is primarily based on three algorithms.
\begin{itemize}
\item{The \emph{selection} policy recursively traverses the tree, selecting good-looking moves for exploration.}
\item{The \emph{simulation} policy randomly or semi-randomly ``plays out'' the game from the current position until a final position.}
\item{The \emph{update} policy modifies the win/loss statistics of nodes based on the simulation results.}
\end{itemize}
Figure \ref{fig:tree} shows a partial Monte-Carlo search tree for the given position. In the example, the selection policy visited the node for \texttt{Black E4}, then \texttt{White A1}. The simulation policy ran, playing out the board with \texttt{E4} occupied by \black\ and \texttt{A1} occupied by \white. This resulted in a loss for \black. The update policy choose to update all nodes along the visited path, incrementing the wins for \white\ moves and incrementing the losses for \black\ moves.

@@

\subsection{The Hivemind System}
Hivemind was designed for experiments in learning to play abstract strategy games on a regular board. The system consists of three modules: the evolutionary learning process, a Monte-Carlo based UCT search tree implementation, and a ``game tracker interface''. The tracker interface is currently implemented for the games of Hex, Go, and Tic-Tac-Toe. It is designed to be possible to implement many different games in the tracker, abstracting away the learning and search process. Hivemind is open-source software, available at \url{http://github.com/etherealmachine/hivemind}.

@@

\subsubsection{Search}
The search module is a simple generic implementation of Monte-Carlo Tree Search, using the UCT algorithm to traverse the tree. Section \ref{mcts} has a description of the MCTS process. The search module uses the Game Tracker to expand the tree and perform the random playouts. The entire source code for search is only about 600 lines of code. Using some extra information from the Game Tracker, search is configurable to use RAVE and AMAF heuristics so a single playout can provide information about the value of multiple nodes.

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{c | c c c}
		 & default & uniform & uniform (tenuki) \\
		\hline
		uniform & 63.50\% & & \\
		uniform (tenuki) & 64.00\% & 50.00\% & \\
		learned & 79.50\% & 81.50\% & 76.00\% \\
		\end{tabular}
	\end{center}
	\caption{All-play-all tournament of the 4 Hivemind variants.}
	\label{fig:results}
\end{figure}
% TODO: run tests using best instead of combined

>>>>

\begin{figure}
	\begin{HexBoard}[board size=13]
		\HGame{c4,f6,e11,i8,h8,i1,i6,d5,i3,h9,h5,h4,i4,g9,f9,g8,f7,f8,e8}
	\end{HexBoard}
	\caption{13x13 Hex Game, \white{} to play.}
	\label{fig:13x13inprogress}
\end{figure}

@@

As stated, the current top computer Hex player, MoHex, uses the open-source Fuego implementation of Monte-Carlo Tree Search (MCTS)\cite{mohex}. The basic Monte-Carlo technique for Computer Go was proposed by Bru\"{u}gmann\cite{brugmann1993monte}. His program \emph{Mango} effectively used a ``flat'' search tree (1-ply) and simply ran simulations from each candidate move to estimate its value. To contrast, MCTS uses Monte-Carlo simulations to grow a minimax search tree. The UCT algorithm was developed alongside MCTS to direct the growth of the search tree\cite{gelly2006exploration}. At any point, the search can be stopped and the current state of the tree used to select a move to play. Performance of the search is primarily based on three algorithms.
\begin{itemize}
\item{The \emph{selection} policy recursively traverses the tree, selecting good-looking moves for exploration.}
\item{The \emph{simulation} policy randomly or semi-randomly ``plays out'' the game from the current position until a final position.}
\item{The \emph{update} policy modifies the win/loss statistics of nodes based on the simulation results.}
\end{itemize}
Figure \ref{fig:tree} shows a partial Monte-Carlo search tree for the given position. In the example, the selection policy visited the node for \texttt{Black E4}, then \texttt{White A1}. The simulation policy ran, playing out the board with \texttt{E4} occupied by \black\ and \texttt{A1} occupied by \white. This resulted in a loss for \black. The update policy choose to update all nodes along the visited path, incrementing the wins for \white\ moves and incrementing the losses for \black\ moves.

@@

\subsection{The Hivemind System}
Hivemind was designed for experiments in learning to play abstract strategy games on a regular board. The system consists of three modules: the evolutionary learning process, a Monte-Carlo based UCT search tree implementation, and a ``game tracker interface''. The design was influenced by Fuego, although Hivemind is written in a different language (Go). In particular, the tracker interface is very similar to Fuego's subclassing system for implementing multiple games\cite{Fuego}. Hivemind's tracker interface is currently implemented for the games of Hex, Go, and Tic-Tac-Toe. The major contribution of Hivemind is the integration of evolutionary learning into the search process. Hivemind is also open-source software, available at \url{http://github.com/etherealmachine/hivemind}.

@@

\subsubsection{Search}
The search module is a simple generic implementation of Monte-Carlo Tree Search, using the UCT algorithm to traverse the tree. Section \ref{mcts} has a description of the MCTS process. The search module uses the Game Tracker to expand the tree and perform the random playouts. The entire source code for search is only about 600 lines of code. Using some extra information from the Game Tracker, search is configurable to use RAVE and AMAF heuristics to update multiple nodes using the results of a single playout.

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{c | c c c}
		 & default & uniform & uniform (tenuki) \\
		\hline
		uniform & 63.50\% & & \\
		uniform (tenuki) & 64.00\% & 50.00\% & \\
		learned & 79.50\% & 81.50\% & 76.00\% \\
		\end{tabular}
	\end{center}
	\caption{All-play-all tournament of the 4 Hivemind variants.}
	\label{fig:results}
\end{figure}
